Title here

### 2016-11-23 portfolio/text/ms.Rmd
Add refs

Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult. These professions tend to have higher year-to-year variability compared to all others (@sethi2012b ; https://www.ers.usda.gov/webdocs/publications/aer812/31365_aer812f.pdf). In addition to economic risk, such as changes in prices, demand, or business cycles, individuals dependent on natural resources are also subject to biological and environmental variability. Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry (http://www.pnas.org/content/104/50/19697.full), or fluctuating prices [include mineral/gas example – Eric can’t find a good one].

Individuals who rely on natural resources for income develop strategies to reduce this variability in year to year earnings. For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends (http://www.nrcresearchpress.com/doi/abs/10.4141/cjps88-001#.WDUQ5aIrLBI). Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.

In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the basket of actions available to individuals to reduce income variability, or incentivize against diversification. In other words, while diversification is generally seen as reducing variability, many individuals don’t. For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize (dairy, grains, soybeans,  http://ageconsearch.umn.edu/bitstream/33967/1/ae020812.pdf). In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004]. In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization (Hilborn et al. 2003 State of the World’s Fisheries).

@@

One of the most well studied examples of diversification are the state and federal fisheries in Alaska. These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources. Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly  $2 billion annually (NMFS 2016) [link]( http://www.st.nmfs.noaa.gov/Assets/commercial/fus/fus15/documents/FUS2015.pdf). The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews. Like much of the rest of the world (Hilborn et al. 2003 state of the world’s fisheries), the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries. And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.

Using a unique dataset collected over 30 years (1985-2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification. The benefit of using individuals rather than vessels is that we are able to track fishes that switch fishery sectors over time, or alter strategies and utilize multiple vessels. We developing and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014]. To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.

>>>>

Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult. These professions tend to have higher year-to-year variability compared to all others [@mishra2002; @sethi2012b]. In addition to economic risk, such as changes in prices, demand, or business cycles, individuals dependent on natural resources are also subject to biological and environmental variability. Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry [@kirilenko2007], or fluctuating prices [include mineral/gas example – Eric can’t find a good one].

Individuals who rely on natural resources for income develop strategies to reduce this variability in year to year earnings. For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends [@zentner1988]. Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.

In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the basket of actions available to individuals to reduce income variability, or incentivize against diversification. In other words, while diversification is generally seen as reducing variability, many individuals don’t. For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize [e.g. dairy, grains, soybeans;  @mishra2002]. In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004]. In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization [@hilborn2003a].

@@

One of the most well-studied examples of diversification are the state and federal fisheries in Alaska. These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources. Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly 2 billion USD annually [@nmfs2016]. The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews. Like much of the rest of the world [@hilborn2003a], the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries. And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.

Using a unique dataset collected over 30 years (1985--2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification. The benefit of using individuals rather than vessels is that we are able to track fishes that switch fishery sectors over time, or alter strategies and utilize multiple vessels. We developing and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014]. To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.
